Untargeted control doc

### 2023-10-12 untargeted/paruntargeted_control.Rmd
Now searching for isos in standards and pooled as well
Allows quantification

# polarity <- "pos"
polarity <- "neg"
# amendment <- "Amm"
# amendment <- "Nit"
# amendment <- "GMP"
amendment <- "Arg"
IS_type <- "noIS"
# IS_type <- "wIS"

@@

mz_group <- function(mz_vals, ppm){
  group_vec <- numeric(length(mz_vals))
  group_num <- 1L
  init_vec <- mz_vals
  names(init_vec) <- seq_along(init_vec)
  while(length(init_vec)>0){
    mz_i <- init_vec[1]
    err <- mz_i*ppm/1000000
    mz_idxs <- init_vec>mz_i-err & init_vec<mz_i+err
    group_vec[as.numeric(names(mz_idxs)[mz_idxs])] <- group_num
    init_vec <- init_vec[!mz_idxs]
    group_num <- group_num+1L
  }
  group_vec
}
```

@@

iso_mz_bounds <- iso_mzs %>%
  left_join(peak_rt_bounds, relationship = "many-to-many") %>%
  left_join(parametadata) %>%
  filter(samp_type=="Smp") %>%
  select(compound_name, feature, n_count, labeled_mzmed, filename, rtmin, rtmax)
iso_areas <- iso_mz_bounds %>%
  split(.$filename) %>% 
  imap(function(file_peaks, file_name_i){
    msdata <- open_dataset("tmzMLs/pqds") %>%
      filter(amendment==!!amendment) %>%
      filter(polarity==!!polarity) %>%
      filter(filename==file_name_i) %>%
      dplyr::collect()
    file_peaks %>%
      pmap(function(...){
        row_data <- data.frame(...)
        area <- msdata[mz%between%pmppm(row_data$labeled_mzmed, 10)] %>%
          .[rt%between%c(row_data$rtmin/60, row_data$rtmax/60)] %>%
          .[,sum(int)]
        cbind(row_data, area=area)
      }) %>%
      bind_rows()
  }, .progress=TRUE) %>%
  bind_rows() %>%
  mutate(labeled_n_count=n_count-1) %>%
  select(compound_name, filename, area, labeled_n_count)
output_csv_write(iso_areas)
```

@@

```{r load isotope mzs and areas}
all_stans <- read_csv("metadata/all_stans.csv") %>%
  filter(polarity==!!polarity) %>%
  filter(compound_type!="Internal Standard") %>%
  select(compound_name, formula)
manual_anno <- readxl::read_excel("untargeted/manual_annotations.xlsx") %>%
  filter(polarity==!!polarity) %>%
  filter(amendment==!!amendment) %>%
  select(feature, compound_name)
best_feat_data <- output_folder %>%
  paste0("best_feat_data.csv") %>%
  read_csv() %>%
  left_join(manual_anno) %>%
  left_join(all_stans)
iso_mzs <- read_csv(paste0(output_folder, "iso_mzs.csv"))
iso_areas <- read_csv(paste0(output_folder, "iso_areas.csv"))
```

>>>>

polarity <- "pos"
# polarity <- "neg"
amendment <- "Amm"
# amendment <- "Nit"
# amendment <- "GMP"
# amendment <- "Arg"
IS_type <- "noIS"
# IS_type <- "wIS"

@@

mz_group <- function(mz_vals, ppm){
  group_vec <- numeric(length(mz_vals))
  group_num <- 1L
  init_vec <- mz_vals
  names(init_vec) <- seq_along(init_vec)
  while(length(init_vec)>0){
    mz_i <- init_vec[1]
    err <- mz_i*ppm/1000000
    mz_idxs <- init_vec>mz_i-err & init_vec<mz_i+err
    group_vec[as.numeric(names(mz_idxs)[mz_idxs])] <- group_num
    init_vec <- init_vec[!mz_idxs]
    group_num <- group_num+1L
  }
  group_vec
}
trapz <- function(x, y) {
  y <- y[order(x)]
  x <- x[order(x)]
  m <- length(x)
  xp <- c(x, x[m:1])
  yp <- c(numeric(m), y[m:1])
  n <- 2*m
  p1 <- sum(xp[1:(n-1)]*yp[2:n]) + xp[n]*yp[1]
  p2 <- sum(xp[2:n]*yp[1:(n-1)]) + xp[1]*yp[n]

  return(0.5*(p1-p2))
}
```

@@

iso_mz_bounds <- iso_mzs %>%
  left_join(peak_rt_bounds, relationship = "many-to-many") %>%
  left_join(parametadata) %>%
  # filter(samp_type=="Smp") %>%
  select(compound_name, feature, n_count, labeled_mzmed, filename, rtmin, rtmax)
iso_areas <- iso_mz_bounds %>%
  split(.$filename) %>% 
  imap(function(file_peaks, file_name_i){
    msdata <- open_dataset("tmzMLs/pqds") %>%
      filter(amendment==!!amendment) %>%
      filter(polarity==!!polarity) %>%
      filter(filename==file_name_i) %>%
      dplyr::collect()
    file_peaks %>%
      pmap(function(...){
        row_data <- data.frame(...)
        area <- msdata[mz%between%pmppm(row_data$labeled_mzmed, 10)] %>%
          .[rt%between%c(row_data$rtmin/60, row_data$rtmax/60)] %>%
          .[,sum(int)]
        cbind(row_data, area=area)
      }) %>%
      bind_rows()
  }, .progress=TRUE) %>%
  bind_rows() %>%
  mutate(labeled_n_count=n_count-1) %>%
  left_join(parametadata) %>%
  select(filename, samp_type, compound_name, labeled_n_count, area)
output_csv_write(iso_areas)
```

@@

```{r load isotope mzs and areas}
all_stans <- read_csv("metadata/all_stans.csv") %>%
  filter(polarity==!!polarity) %>%
  filter(compound_type!="Internal Standard") %>%
  select(compound_name, formula)
manual_anno <- readxl::read_excel("untargeted/manual_annotations.xlsx") %>%
  filter(polarity==!!polarity) %>%
  filter(amendment==!!amendment) %>%
  select(feature, compound_name)
best_feat_data <- output_folder %>%
  paste0("best_feat_data.csv") %>%
  read_csv() %>%
  left_join(manual_anno) %>%
  left_join(all_stans)
iso_mzs <- read_csv(paste0(output_folder, "iso_mzs.csv"))
iso_areas <- read_csv(paste0(output_folder, "iso_areas.csv")) %>%
  filter(samp_type=="Smp")
```
